---
format: html
editor: 
  markdown: 
    wrap: 72
---

# Getting Started in R

#### Instructions for installing R and R Studio: <https://posit.co/download/rstudio-desktop/>.

### Additional Resources

1.[Big Book of R](https://www.bigbookofr.com/)

2\. Learning R in R using package
[swirl](https://swirlstats.com/students.html) Helpful website for asking
questions: [Stackoverflow](https://stackoverflow.com)

3\. [Learning Statistics with R](https://learningstatisticswithr.com/)
(free access!)

4\. [Data wrangling in R: load, explore, transform, and visualize data
for modeling tidyverse
libraries](https://learning.oreilly.com/library/view/data-wrangling-with/9781803235400/)
(UNH students/faculty can access via the UNH Library)

5\. [Guide to multilevel
modeling](https://www.learn-mlms.com/index.html)

6\. [Guide for poisson
regressions](https://stats.oarc.ucla.edu/r/dae/poisson-regression/)

7\. [Choosing a graph type (data to viz)](https://www.data-to-viz.com/)

8\. [Guide for making grahs (this resource is linked to datatoviz in
#6!)](https://r-graph-gallery.com/)

9\. [Guide for choosing
models](https://pmc.ncbi.nlm.nih.gov/articles/PMC8187274/#ecy3336-sec-0004)

10\. [Guide to creating pdf documents using
Quarto](https://quarto.org/docs/guide/)

11\. [Guide to publishing manuscripts using
Quarto](https://quarto.org/docs/manuscripts/)

12\. [Guide for formatting markdown
documents](https://rmd4sci.njtierney.com/)

# Schedule & Materials

```{r}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE
library(googlesheets4)
library(gt)
library(tidyverse)
library(fontawesome)
gs4_deauth()
d <- read_sheet("https://docs.google.com/spreadsheets/d/1k4RPfWwKVxPpp4wRDEc08KDxzBLG3vOcWMC6_hsF648/edit?gid=0#gid=0", sheet = "Spring2025")
```

```{r, echo = FALSE}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE
library(googlesheets4)
library(gt)
library(tidyverse)
library(fontawesome)

# Deauthorize Google Sheets for reading public sheets
gs4_deauth()

# Read the data from the Google Sheet
d <- read_sheet("https://docs.google.com/spreadsheets/d/1k4RPfWwKVxPpp4wRDEc08KDxzBLG3vOcWMC6_hsF648/edit?gid=0", sheet = "Spring2025")

# Render the table
d |>
  gt() |>
  
  # Handle missing values in the first two columns
  sub_missing(
    columns = c(1:2),
    missing_text = ""
  ) |>
  
  # Add file-code icon with links for the 'Files' column
  text_transform(cells_body(columns = c(Files)),
    fn = function(x) {
      ifelse(is.na(x), fontawesome::fa("file-code", fill_opacity = 0.1),
        glue::glue("<a href = '{x}' >{fontawesome::fa('file-code')}</a>")
      )
    }
  ) |>
  
  # Add table icon with links for the 'Data' column
  text_transform(cells_body(columns = c(Data)),
     fn = function(x) {
        ifelse(is.na(x), fontawesome::fa("table", fill_opacity = 0.1),
          glue::glue("<a href = '{x}'>{fontawesome::fa('table')}</a>")
        )
     }
   ) |>
  
  # Add link icon with links for the 'Videos' column
  text_transform(cells_body(columns = c(Videos)),
    fn = function(x) {
       ifelse(is.na(x), fontawesome::fa("link", fill_opacity = 0.1),
         glue::glue("<a href = '{x}'>{fontawesome::fa('link')}</a>")
       )
    }
  ) |>
  
  # Align columns: center for icons, left for dates/topics
  cols_align(
    align = "center", 
    columns = c(Files, Data, Videos)
  ) |>
  
  # Align Date and Topic columns to the left
  cols_align(
    align = "left", 
    columns = c(Date, Topic)
  ) |>
  
  # Add a top border to non-empty rows
  tab_style(
    style = cell_borders(
      sides = "top",
      color = "black", 
      weight = px(2)
    ),
    locations = cells_body(rows = !is.na(Date))
  ) |>
  
  # Add a thick bottom border after a specific date (customize as needed)
  tab_style(
    style = cell_borders(
      sides = "bottom",
      color = "black",
      weight = px(4)
    ),
    locations = cells_body(rows = Date == "09/09/2024") # Adjust date if needed
  ) |>
  
  # Add a border around the column labels
  tab_style(
    style = cell_borders(
      sides = c("top", "bottom"),
      color = "black",
      weight = px(4)
    ),
    locations = cells_column_labels(columns = everything())
  ) |>
  
  # Set custom column widths for better alignment
  cols_width(
    Date ~ px(250),
    Topic ~ px(250),
    Files ~ px(100),
    Data ~ px(100),
    Videos ~ px(100)
  )|>
  # Set table width to span a % of the available page width
  tab_options(
    table.width = pct(70)
  )
```
