---
title: "Untitled"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---
```{r, include = FALSE}
library(tidyverse)
library(palmerpenguins)
```

## Assumptions of Linear Regressions


$y = mx + b$

```{r}
str(penguins)
summary(penguins)


ggplot(aes(x = body_mass_g, y = flipper_length_mm), data = penguins) +
  geom_point() +
  geom_smooth(method = "lm")


mod <- lm(body_mass_g ~ flipper_length_mm + species, data = penguins)
summary(mod)

confint(mod, level = 0.90)

penguins_short <- penguins %>% 
  select(-c(bill_length_mm,bill_depth_mm,flipper_length_mm))
str(penguins_short)

mod2 <- lm(body_mass_g ~., data = penguins_short)
summary(mod2)

mod3 <- lm(body_mass_g ~.-island, data = penguins_short)
summary(mod3)
```

## Plot diagnostics

```{r}
library(broom)
par(mfrow = c(2,2))
plot(mod3)

mod3fit %>% 
  select(-c(.rownames, .sigma)) %>% 
  top_n(3, wt = .cooksd)
#observation has high influence if Cook's distance exceeds 4/(n - p - 1)
#n = number of observations
#p = number of predictors 
cook <- 4/(nrow(penguins_short) - 2)
```
-  residuals vs fitted: used to check linear relationship assumption. A horizontal line without a distinct pattern is wanted. 
-  Q-Q residuals: used to check normal distribution. It is good if the lines follow the straight dashed line
-  Scale location: used to check homogeneity of variance of residuals. Horizontal line with equally dispersed points is good. 
-  Residuals versus leverage. Used to identify outliers that can influence the model. 


